Bandwidth-Aware Reconfigurable Memory Design with Hybrid Memory Technologies

### 2011-04-09 paper.tex

The bandwidth problem of multicore processors has drawn much attention recently.
Yu \emph{et al.} proposed a LLC partitioning algorithm to minimize bandwidth
requirement to off-chip main memory~\cite{Yu:2010:OCMB}. While most cache
partitioning techniques focus on cache miss rates, their work takes a different
approach in which tasks¡¯ memory bandwidth requirements are taken into account
when identifying a cache partitioning for multi-programmed and/or multithreaded
workloads. Cache resources are allocated with the objective that the overall
system bandwidth requirement is minimized for the target workload. The key
insight is that cache miss-rate information may severely misrepresent the actual
bandwidth demand of the task, which ultimately determines the overall system
performance and power consumption. <the moguls paper, the figure 1, however design for
induvidual application, i.e., with different application can generate different
designs.>

A large body of recent research focuses on exploring new memory technologies to
trade off between latency, bandwidth, and cost. Wu et al. explored various
memory technologies - SRAM, eDRAM [23], magnetic RAM (MRAM) [12] and PCM to best
construct L3 caches in terms of performance and power [24]. The eDRAM and MRAM
technologies are evaluated as LLCs by stacking each memory on top of the
processor die. Both studies mainly focused on reducing the latency gap between L2
cache~(LLC) and external memory and do not examine the bandwidth. <nvm cache
research: most for latency, not bandwidth opt.>

>>>>

The bandwidth problem of multicore processors has drawn much attention recently.
Yu \emph{et al.} proposed a LLC partitioning algorithm to minimize bandwidth
requirement to off-chip main memory~\cite{Yu:2010:OCMB}. While most cache
partitioning techniques focus on cache miss rates, their work takes a different
approach in which tasks¡¯ memory bandwidth requirements are taken into account
when identifying a cache partitioning for multi-programmed and/or multithreaded
workloads. Cache resources are allocated with the objective that the overall
system bandwidth requirement is minimized for the target workload. The key
insight is that cache miss-rate information may severely misrepresent the actual
bandwidth demand of the task, which ultimately determines the overall system
performance and power consumption. However, their work only focused on LLC (L2
cache). Our design targe is the overall on-chip memory system, which provides
more design dimensions and flexibility. An analytical performance model was
proposed by Sun et al. to explore the design space of memory hierarchies for
throughput computing [moguls]. However, their memory hierarchy design was
tailored for individual applications, which was undesirable for a real system
design.

A large body of recent research focuses on exploring new memory technologies to
trade off between latency, bandwidth, and cost. Wu et al. explored various
memory technologies - SRAM, eDRAM [23], magnetic RAM (MRAM) [12] and PCM to best
construct L3 caches in terms of performance and power [24]. The eDRAM and MRAM
technologies are evaluated as LLCs by stacking each memory on top of the
processor die. Both studies mainly focused on reducing the latency gap between
L2 cache~(LLC) and external memory, but do not examine the design from bandwidth
perspective.
